---
title: "Homework 2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1
Read and clean the data
```{r}
subway_data = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```
The dataset contains line, station name, station latitude and longitude, routes served, entry, vending, entrance type and ADA compliance. My data cleaning steps so far include: load the data and clean up the column names, select columns that I want to keep, and convert the entry variable from character to a logical variable.
The dimension (rows x columns) of the resulting dataset is `r dim(subway_data)`.
These data are not tidy: the route number is spread across 11 columns.

Answer the following questions using these data:

```{r}
distinct_data = distinct(subway_data, line, station_name, .keep_all = TRUE)
```
How many distinct stations are there?

There are `r nrow(distinct_data)` distinct stations.

How many stations are ADA compliant?

There are `r sum(distinct_data$ada, ada = TRUE)` stations are ADA compliant.

What proportion of station entrances / exits without vending allow entrance?
```{r}
distinct_data = mutate(distinct_data, vending = recode(vending, "YES" = TRUE, "NO" = FALSE))
```
The proportion of station entrances / exits without vending allow entrance is 
`r 1 - sum(distinct_data$vending, na.rm = FALSE) / length(distinct_data$vending)`

Reformat data so that route number and route name are distinct variables. 
```{r}
distinct_tidy = gather(distinct_data, key = route_number, value = route_name, 
                       route1:route11, na.rm = TRUE)
```
How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

`r nrow(filter(distinct_tidy, route_name == "A"))` distinct stations serve the A train.

Of the stations that serve the A train, `r nrow(filter(distinct_tidy, route_name == "A", ada == TRUE))` are ADA compliant.

## Problem 2
Read and clean the Mr.Trash Wheel sheet:
```{r}
library(readxl)
library(cellranger)
mr_trash_data = 
  read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 1,
             range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster), month != "Grand Total") %>% 
  mutate(sports_balls = round(sports_balls), 
         sports_balls = as.integer(sports_balls))
```

Read and clean precipitation data for 2016 and 2017:
```{r}
precipitation_2016 =
  read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 4,
             range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = "2016")

precipitation_2017 =
  read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 3,
             range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = "2017")
```

Combine datasets and convert month to a character variable:
```{r}
precipitation_data = 
  bind_rows(precipitation_2016, precipitation_2017) %>% 
  mutate(month = month.name[month])
```
The number of observations in "mr_trash_data" is `r nrow(mr_trash_data)`.
The number of observations in "precipitation_data" is `r nrow(precipitation_data)`.

Examples of key variables:

The total precipitation in 2017 is `r sum(precipitation_2017$total)`.
The median number of sports balls in a dumpster in 2016 is 
`r median(subset(mr_trash_data, year == 2016)$sports_balls)`.