---
title: "Homework 2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1
Read and clean the data
```{r}
subway_data = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```
The dataset contains line, station name, station latitude and longitude, routes served, entry, vending, entrance type and ADA compliance. My data cleaning steps so far include: load the data and clean up the column names, select columns that I want to keep, and convert the entry variable from character to a logical variable.
The dimension (rows x columns) of the resulting dataset is `r dim(subway_data)`.
These data are not tidy: the route number is spread across 11 columns.

Answer the following questions using these data:

```{r}
distinct_data = distinct(subway_data, line, station_name, .keep_all = TRUE)
```
How many distinct stations are there?

There are `r nrow(distinct_data)` distinct stations.

How many stations are ADA compliant?

There are `r sum(distinct_data$ada, ada = TRUE)` stations are ADA compliant.

What proportion of station entrances / exits without vending allow entrance?
```{r}
distinct_data = mutate(distinct_data, vending = recode(vending, "YES" = TRUE, "NO" = FALSE))
```
The proportion of station entrances / exits without vending allow entrance is 
`r 1 - sum(distinct_data$vending, na.rm = FALSE) / length(distinct_data$vending)`